A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-26 latex/paper.tex
Worked in Section II

% Dask
On the other hand, Dask was created with the purpose of natively parallelize Python
pipelines while keeping the syntax of familiar API from the Python's scientific
ecosystem. However, Dask is still a young framework with work to be done; it's API
does not completely replicate the library it supports. While its lower-level API
allows the implementation of more complex algorithms it sacrifices a layer of
optimization.
% Related work
Previous work shows that Dask had significant overhead and was
hard to debug~\cite{Mehta:17}.
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework.
Therefore we think it is valuable to re-compare it with Spark.

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality, and lazy
evaluation. Spark's main data structure, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
elements. 
Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
performance features. 
Spark's core is written in Scala, but it also has an R and a Python API
called PySpark. We used PySpark, given the popularity of Python in
neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{yarn} and
Mesos~\cite{mesos} are also available. In the Spark standalone scheduler, a
process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches
tasks to them. The Spark standalone scheduler uses a LIFO
(Last-In-First-Out) task scheduling policy \TG{we should describe it
briefly and refer to the paper}. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated  process
on the cluster, and cluster mode, where the driver runs in a worker. We
used client mode as PySpark is not available in cluster mode. We used
Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. \TG{A few more
introductory sentences should be added, following Spark's model.}
Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler.
Dask Distributed also has a LIFO policy. In theory, it should speed-up the
pipeline execution by finishing a branch of task computations before
starting new ones \TG{This is meant to reduce the memory footprint of the
application}. \TG{Here we should add a few comparison sentences between
Spark's and Dask's schedulers.} Dask has five main APIs and data structures: Bag,
Array, Futures, Delayed and DataFrame. We used them all
except DataFrame as it is not adapted to the type of data we
used. All APIs provide in-memory computing, data locality, and lazy
evaluation, however, none of them has fault tolerance \TG{Check fault
tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a
programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}.
An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for
the processing of large arrays. It provides a distributed clone of the
popular
NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
tasks that do not fit in the Array, DataFrame or Bag frameworks. 
Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
Delayed as they support arbitrary tasks, but they operate in real-time rather
than lazily.

@@

We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
brain. The original data is stored in 125 blocks in the MINC~\cite{minc}
HDF5-based format, available at
\url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
resolution of \SI{40}{\micro\metre}. We converted the blocks into the
\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
a total data size of 76~GiB \TG{Check if GiB or GB}. 
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
2.5GB, 0.6GB, and
0.1GB \TG{check sizes}, using the sam~\cite{sam} library.

@@

\subsection{Applications}
%% Incrementation
\subsubsection{\textbf{Incrementation}}
As our first experiment to benchmark the frameworks, we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It reads blocks from the BigBrain
image, increments the intensity value of each voxel by 1 for a specified amount of
iterations, and writes the result to an NFS as a NifTI image. Incrementing the data
reduce the caching effect. This experiment allows us to study the behavior of the
frameworks when all inputs are processed independently; i.e.\ each task in the graph
only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimics the
behavior of analyzing multiple independent subjects in parallel.

This experiment refers to the implementation of the Spark pipelines purposed
in Ref.~\cite{Hayot-Sasson:17}.

\begin{algorithm}[!b]
    \caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a block}
    \Require{\(fs\), NFS to write image to.}
    \State{Read \(block\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(block \in image\)}
            \State{\(block\gets block+1\)}
            \State{Sleep \(x\)}
        \EndFor
    \EndFor
    \State{Write \(block\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation}\label{fig:tg-inc}
\end{figure}

\subsubsection{\textbf{Histogram}}
Our second experiment consists to calculate the histogram of the intensity of the
voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
NFS, calculate individual image frequency, aggregate the frequency of the images
together, and write the result to a file. In this experiment, the result depends on
all the input (see Fig.~\ref{fig:tg-histo}). This allows the study of the behavior of
the frameworks when tasks are dependent. Also, this application requires data
shuffling thus inter-worker communication. We decided to omit Dask Futures as it does
not have any value over Dask Delayed in this application.

@@

\subsection{Experiment}
The applications are benchmarked according to up to four parameters
(Table~\ref{tab:param}):

\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
scaling of the frameworks in a cluster. Also, it helps in determining the level of
the parallelism of the frameworks.

\textbf{Number of blocks:} Evaluating the frameworks on a different amount of blocks
allows us to understand their behavior with different task size and IO requirements.

\textbf{Number of iterations:} We change the number of iterations for the
incrementation~\ref{alg:incrementation} to study the effect of the job length.

\textbf{Sleep delay:} We modify the sleep delay in the
incrementation~\ref{alg:incrementation} to understand the behavior of different task
lengths.

The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
seconds sleep delay.

@@

\subsection{Infrastructure}
All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
7.5.1804 as the base operating system and have the Linux kernel version
3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
30GB 2666 MHz memory. The network has a 10GB/s bandwidth. The filesystem is 2TB NFS
v4.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}

%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of instances}

>>>>

% Dask
On the other hand, Dask was created with the purpose of natively parallelize Python
pipelines while keeping the syntax of familiar API from the Python's scientific
ecosystem. However, Dask is still a young framework with work to be done; it's API
does not completely replicate the library it supports. While its lower-level API
allows the implementation of more complex algorithms it sacrifices a layer of
optimization.
% Related work
Previous work shows that Dask had significant overhead and was
hard to debug~\cite{Mehta:17}.
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework.
Therefore we think it is valuable to re-compare it with Spark.
\TG{Cite~\cite{hayot2019performance}}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
engine. It provides in-memory computing, data locality, and lazy
evaluation. Spark's main data structure, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
elements. 
Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
performance features. 
Spark's core is written in Scala, but it also has an R and a Python API
called PySpark. We used PySpark, given the popularity of Python in
neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler, a
process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches
tasks to them. The Spark standalone scheduler uses a LIFO
(Last-In-First-Out) task scheduling policy \TG{we should describe it
briefly and refer to the paper}. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated  process
on the cluster, and cluster mode, where the driver runs in a worker. We
used client mode as PySpark is not available in cluster mode. We used
Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. \TG{A few more
introductory sentences should be added, following Spark's model.}
\TG{Explain workers}
Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler.
Dask Distributed also has a LIFO policy. In theory, it should speed-up the
pipeline execution by finishing a branch of task computations before
starting new ones \TG{This is meant to reduce the memory footprint of the
application}. \TG{Here we should add a few comparison sentences between
Spark's and Dask's schedulers.} Dask has five main APIs and data structures: Bag,
Array, Futures, Delayed and DataFrame. We used them all
except DataFrame as it is not adapted to the type of data we
used. All APIs provide in-memory computing, data locality, and lazy
evaluation, however, none of them has fault tolerance \TG{Check fault
tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a
programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}.
An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for
the processing of large arrays. It provides a distributed clone of the
popular
NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
tasks that do not fit in the Array, DataFrame or Bag APIs. 
Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
Delayed as they support arbitrary tasks, but they operate in real-time rather
than lazily.

\subsection{Infrastructure}

 We used Compute Canada's
 \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud}
 operated by the \href{https://www.westgrid.ca}{WestGrid} regional
 organization at the University of Victoria, and running OpenStack
 \TG{version?}. We used c8-30gb-186 cloud instances with 8 VCPUs, an Intel
 Xeon Gold 6130 processor, 30~GB of RAM at 2666~MHz, 20~GB of mounted
 storage, and a base image running CentOS 7.5.1804 with  Linux kernel
 version
 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a 10Gb/s Ethernet network.

 Cloud instances hosted a single Dask or Spark worker, configured to use 8
 threads. We used Dask's default configuration that uses all the available
 memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
 configured Spark to use 1 executor per worker and 25~GB of memory per
 executor, to leave 5~GB for overhead. \TG{Check spill to disk}

 One cloud instance did not host any worker, and had a 2~TB disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. We configured the Spark driver to use 25~GB of memory, and
 used the default configuration for the master.

@@

We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
brain with voxel intensities ranging from 0 to 65,535. The original data is
stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
resolution of \SI{40}{\micro\metre}. We converted the blocks into the
\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
a total data size of 76~GiB \TG{Check if GiB or GB}. 
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
2.5GB, 0.6GB, and
0.1GB \TG{check sizes}, using the sam~\cite{sam} library.

@@

\subsection{Applications}

We used three neuroimaging applications to evaluate the engines in
different conditions. The first two ones, incrementation and histogram, are
simple synthetic benchmarks representing basic map-only and map-reduce applications.
The third one is a more realistic application representative of popular
BIDS applications~\cite{gorgolewski2017bids}.

%% Incrementation
\subsubsection{Incrementation}
We used the simple image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate a more complex
processing, repeats this process for a specified amount of iterations, and
finally writes the result as a NifTI image back to the shared file system.
This application allows us to study the behavior of the engines when all
inputs are processed independently, in a map-only way (see
Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
independent subjects in parallel.

\begin{algorithm}[!b]
    \caption{Incrementation (adapted from~\cite{hayot2019performance})}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a block}
    \Require{\(fs\), NFS to write image to.}
    \State{Read \(block\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(block \in image\)}
            \State{\(block\gets block+1\)}
            \State{Sleep \(x\)}
        \EndFor
    \EndFor
    \State{Write \(block\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations, represented
    for 3 BigBrain blocks. Circles represent the incrementation and sleep
    function while rectangles represent image blocks.}\label{fig:tg-inc}
\end{figure}

\subsubsection{Histogram}

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
 file system, calculates intensity frequencies, aggregates the frequencies
 across the blocks, and finally writes the resulting histogram to the
 shared file system as a single \TG{X}KB file. This application implements
 a typical map-reduce pattern, where the final result is obtained from all
 the input blocks (see Fig.~\ref{fig:tg-histo}). This application requires
 data shuffling thus inter-worker communication. The total amount of shuffled
 data is however limited to \TG{X}MB as it only consists of image
 histograms.

 \TG{Add links to the Github repo for applications and any other relevant file.}

@@

\subsubsection{BIDS app example}

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
application available on GitHub. This application operates in a map-reduce
way. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
resulting image (\TG{X}MB on average) to the shared file system. The reduce
phase, also called group analysis, computes the volume of each brain and
returns the average volume, shuffling a total of \TG{X}GB image data.

While Incrementation and Histogram were implemented directly in Python,
BIDS app example requires an external library and is distributed as a
Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
transfers from DockerHub, we preloaded our OpenStack instances with the
Docker image, and started the Docker engine at startup. \TG{explain run.py modifications if relevant.}

\subsection{Experiments}

We varied four parameters in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram to measure the effect of different IO patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation to evaluate the effect of job length.

% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
% seconds sleep delay.

@@

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}
%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of instances}
